Generating a Sierpinski Triangle and a Sierpinski Carpet
========================================================
<script src="ga.js"></script>
<h2>Sierpinki Triangle</h2>

A <a href="http://en.wikipedia.org/wiki/Sierpinski_triangle">Sierpinski Triangle</a> is a fractal triangle that is produced by a very simple algorithm.

- Draw three vertices

- Randomly generate a point on the graph.

- Randomly choose one of the three vertices

- Go the point halfway between your random point and your random vertex and set this as your current location.

- Choose another random vertex and go halfway between your previous current location and the new randomly chosen vertex.

- Do this a large number of times, marking each point you land on, and the below triangle will begin to emege.

The code used is
```{r}
sierpinskiTriangle = function(numPoints = 10000) {
  currentX <- runif(1);
  currentY <- runif(1);
  
  xCorners <- c(0, 1, .5);
  yCorners <- c(0, 0, 3^.5 / 2);
  
  xPoints <- vector();
  yPoints <- vector();
  
  xPoints <- c(xPoints, currentX);
  yPoints <- c(yPoints, currentY);
  
  for(i in 1:numPoints) {
    corner <- randomInt(3);
    currentX <- (currentX + xCorners[corner]) / 2;
    currentY <- (currentY + yCorners[corner]) / 2;
    
    xPoints <- c(xPoints, currentX);
    yPoints <- c(yPoints, currentY);
  }
  plot(xPoints, yPoints, pch=20);
}

randomInt <- function(max) {
  return(ceiling(runif(1) * max));
}
```

If this is saved in a document called "fractal.r" in your working directory, just load it and run it as below:

```{r}
source('fractal.r');
sierpinskiTriangle(50000);
```

If you see a few points that seem out of place, it is because you start with a random point and it may take a few iterations of the algorithm before it falls into the pattern.


<h2>Sierpinski Carpet</h2>

A <a href="http://en.wikipedia.org/wiki/Sierpinski_carpet">Sierpinski Carpet</a> has a similar look, but is a square instead of a triangle.  To make it, you start with a filled square and recursively remove the middle third.  That is, you remove the middle square and recursively remove the middle square of each of the surrounding eight blocks.

The function I wrote is:
```{r}
sierpinskiCarpet <- function(depth = 4, shouldReturn = FALSE) {
  if(depth < 1) {
    stop("can't have a depth less than 1");
  }
  width = 3 ^ depth;
  carpet <- matrix(1, nrow=width, ncol=width);
  carpet[(width/3 + 1):(2 * width/3),(width/3 + 1):(2 * width/3)] <- 0;
  if(depth > 1) {
    for(i in 0:2) {
      for(j in 0:2) {
        if(i != 1 || j != 1) {
        iStart <- 1 + (width / 3) * i;
        iEnd <- iStart + width / 3 - 1;
        jStart <- 1 + (width / 3) * j;
        jEnd <- jStart + width / 3 - 1;
        carpet[iStart:iEnd, jStart:jEnd] <- sierpinskiCarpet(depth - 1, TRUE);
        }
      }
    }
  }
  if(shouldReturn) {
    return(carpet);
  } else {
    image(carpet, col=c("white", "black"));
  }
}
```

Again, if this is in a file called "fractal.r", use the following two lines:

```{r}
source('fractal.r')
sierpinskiCarpet(5, FALSE);
```

<object name="footer" type="text/html" data="footer.html"></object>